Containers and R---the Rockerverse and beyond

### 2019-11-06 manuscript.Rmd

@@

- GPU environments [@noamross, @cboettig]
- [RCloud](https://github.com/att/rcloud/tree/master/docker) ?

>>>>

One use for containers is to run on shared local hardware where teams manage their own high-performance servers.
This can follow one of several design patterns: users may deploy containers to hardware as a work environment for a specific project, conatiners may provide per-user persistent environments, or a single container can act as a common multi-user environment for a server.
In all cases, though, the containerized approach provides several advantages: First, users may use the same image and thus work environment on desktop and laptop computers, as well.
The former models provide modularity, while the latter approach is most similar to a simple shared server.
Second, software updates can be achieved by updating and redeploying the container, rather than tracking local installs on each server.
Third, the containerized environment can be quickly deployed to other hardware, cloud or local, if more resources are neccessary or in case of server destruction or failure.
In any of these cases, users need a method to interact with the containers, be it and IDE, or command-like access and tools such as SSH, which is usually not part of standard container recipes and must be added.
The Rocker project provides containers pre-installed with the RStudio IDE.

In cases where users store nontrivial amounts of data for their projects, data needs to persist beyond the life of the container.
This may be via in shared disks, attached network volumes, or in separate storage where it is uploaded between sessions.
In the case of shared disks or network-attached volumes, care must be taken to persist user permissions, and of course backups are still neccessary.
When working with multiple servers, an automation framework such as [Ansible](https://www.ansible.com) may be useful for managing users, permisions, and disks along with containers.

In some cases, containers are not completely portable between hardware environments.
This is the case for software using graphical processing units (GPUs) which are increasingly popular for compute-intensive machine learning tasks.
Containers running GPU software  require drivers and libraries specific to GPU models and versions, and containers require a specialized runtime to connect to the underlying GPU hardware.
For NVIDIA GPUs, the [NVIDIA Container Toolkit](https://github.com/NVIDIA/nvidia-docker) includes a specialized runtime plugn for docker and set of base images with appropriate drivers and libraries, and the Rocker project has (beta) images based on these that include GPU-enabled versions of machine-learning R packages.

@@

[RCloud](https://github.com/att/rcloud/tree/master/docker) uses a `rocker/drd` base image for creation of collaborative data analysis and visualisation environments.
